---
title: "rediMR_summary"
output: html_document
date: "2024-06-12"
---

```{r setup, include=FALSE, error=F, message=F, out.height="60%", out.width="90%"}
knitr::opts_chunk$set(echo=F, message=F,warning=F, 
                      fig.path="../output/")

# load required packages
suppressMessages(silent <- lapply(c("tidyverse", "data.table", "parallel", "paletteer", "RColorBrewer", "ggpubr"),  
       library, character.only = T))

# Run pantry.R file
source("../scripts/pantry.R")

```

```{r params}

redimrDir="../data/processed/rediMR"
mrDir="../data/processed/MR"

# Define parameters
tag="v2"
foods <- c("raw_veg", "fresh_fruit")

```


```{r} 
# ==========================
## Load all raw_veg data 
# ==========================

veg_tag <- paste0("raw_veg_", tag)

veg_dat = readRDS(paste0(redimrDir, "/", veg_tag, "_datInput.rda"))
veg_dPCs = readRDS(paste0(redimrDir, "/", veg_tag, "_dietPCs.rda"))

## raw_veg rediMR outputs by covariate group 
veg_Bchange.l <- list(
  All <- lapply(covarSetGroups$dietPCs$Sets, function(set) {
    veg_tag_set <- paste0(veg_tag, "_", set)
    list(BchangeAll <- read.csv(paste0(redimrDir, "/", veg_tag_set, "_tabBchangeAllCov.csv"))) }),
  Each <- read.csv(paste0(redimrDir, "/", veg_tag, "_dietpctop23_tabBchangeByCov.csv")))
names(veg_Bchange.l) <- c("All", "By") ; names(veg_Bchange.l$All) <- covarSetGroups$dietPCs$Sets


veg_Bchange.l$All_combined <- do.call(rbind.data.frame, lapply(veg_Bchange.l$All, function(set) { 
    set[[1]] }) ) %>%
    mutate(nPCs=as.numeric(gsub("Top", "", gsub("DietPC.*", "", Covar)))) %>%
    mutate(snp=gsub("snp","",gsub("[.]",":",snp)),
           Covar=factor(Covar, levels=covarSetGroups$dietPCs$Labels)) 
  
  
# raw_veg rediMR outputs for confounders
veg_Bchange_conf.l <- list(
  BchangeAll = read.csv("../data/processed/rediMR/raw_veg_v2_stnd_tabBchangeAllCov.csv"),
  BchangeBy = read.csv("../data/processed/rediMR/raw_veg_v2_stnd_tabBchangeByCov.csv")
)


# ==========================
## Load all fresh_fruit data 
# ==========================

fruit_tag <- paste0("fresh_fruit_", tag)

fruit_dat = readRDS(paste0(redimrDir, "/", fruit_tag, "_datInput.rda"))
fruit_dPCs = readRDS(paste0(redimrDir, "/", fruit_tag, "_dietPCs.rda"))

## fresh_fruit rediMR outputs by covariate group 
fruit_Bchange.l <- list(
  All <- lapply(covarSetGroups$dietPCs$Sets, function(set) {
    fruit_tag_set <- paste0(fruit_tag, "_", set)
    list(BchangeAll <- read.csv(paste0(redimrDir, "/", fruit_tag_set, "_tabBchangeAllCov.csv"))) }),
  Each <- read.csv(paste0(redimrDir, "/", fruit_tag, "_dietpctop23_tabBchangeByCov.csv")))
names(fruit_Bchange.l) <- c("All", "By") ; names(fruit_Bchange.l$All) <- covarSetGroups$dietPCs$Sets


# fresh_fruit rediMR outputs for confounders
fruit_Bchange_conf.l <- list(
  BchangeAll = read.csv("../data/processed/rediMR/fresh_fruit_v2_stnd_tabBchangeAllCov.csv"),
  BchangeBy = read.csv("../data/processed/rediMR/fresh_fruit_v2_stnd_tabBchangeByCov.csv")
)


```


```{r data-prep-2}

# ===================================================================
## Compiling dietPC and confounder adjusted SNP-food associations
# ===================================================================

# Adjusting for ALL dietPCs PLUS EACH behavioral/SES confounders
veg_BchangeEachConf <- readRDS("../data/processed/rediMR/veg_BchangeEachConf.rda")
veg_BchangeAddConf <- readRDS("../data/processed/rediMR/veg_BchangeAddConf.rda")


# Adjusting for ALL dietPCs PLUS ADDITIONAL behavioral/SES confounders
veg_Bdat.l <- list(
  each_dPC = veg_Bchange.l$By,
  each_conf = veg_Bchange_conf.l$BchangeBy,
  each_allvars = rbind.data.frame(veg_Bchange.l$By, veg_Bchange_conf.l$BchangeBy),
  each_dPCplusConf = do.call(rbind.data.frame, veg_BchangeEachConf),
  all_dPCs = veg_Bchange.l$All$dietpctop23[[1]] %>% select(-c("X", "RefinedSet")),
  all_conf = veg_Bchange_conf.l$BchangeAll %>% select(-c("X", "RefinedSet")),
  all_allvars = rbind.data.frame(veg_Bchange.l$All$dietpctop23[[1]], veg_Bchange_conf.l$BchangeAll),
  all_dPCplusConf = do.call(rbind.data.frame, veg_BchangeAddConf)
)

```

#### Check characteristics of diet PCs (derived without the diet trait)

```{r}

# ========
## PVE
# ========

ggarrange(
  plot_dietPC_pve.fun(veg_dPCs$sdev, "dietPCs without raw vegetable"),
  plot_dietPC_pve.fun(fruit_dPCs$sdev, "dietPCs without fruit intake"))

```


```{r}

## Raw vegetables - dietPCs
plot_dietPC.fun("raw_veg", veg_dPCs$rotation, 10) #%>%
  #ggsave(filename="../output/p_veg_dietPCs_dietGroup2.pdf", height=10, width=20)

## Fresh fruit - dietPCs
#plot_dietPC.fun("fresh_fruit", fruit_dPCs$rotation , 10) #%>%
  #ggsave(filename="../output/p_fruit_dietPCs_dietGroup2.pdf", height=10, width=20)

```

Checked assumptions:
-Diet (specific traits & diet patterns) associate with behavioral, SES and biological confounders.
-Phenotypic diet-confounder associations are attenuated, after adjusting for dietPCs


```{r}

library(gplots)
veg_cormat <- readRDS("../data/processed/plots/raw_veg_v2_dietGroup2_plotHMconf.rda")
veg_cormat_dG2 <- veg_cormat[c("raw_veg", covarSets$dietpctop23$Covars),-1]

#pdf(paste0("../output/", veg_tag, "_plotHMconf_dietGroup2.pdf"), height=4, width=7)
heatmap.2(t(veg_cormat_dG2[c("raw_veg", covarSets$dietpctop23$Covars),]),
          col = palettes$hm_bwr,
          margins = c(8,8), #key="False",
          Rowv=T, Colv=F, 
          colsep = 1,
          offsetCol = 0, offsetRow = 0, #lhei=c(1,5), lwid = c(1,6),
          linecol =  "black",
          density.info = "none", trace = "none", dendrogram = "none",
          cexRow = 0.75, cexCol = 0.75)
#dev.off()
       

# ===============
## Fresh fruit
# ===============

# Create correlations matrix of all dietPCs with confounders
#fruit_cormat <- readRDS("../data/processed/plots/fresh_fruit_v2_dietGroup2_plotHMconf.rda")
#fruit_cormat_dG2 <- fruit_cormat[c("fresh_fruit", covarSets$dietpctop23$Covars),-1]

#pdf(paste0("../output/", fruit_tag, "_plotHMconf_dietGroup2.pdf"), height=4, width=7)
#heatmap.2(t(fruit_cormat_dG2[c("fresh_fruit", covarSets$dietpctop23$Covars),]),
#          col = palettes$hm_bwr,
#          margins = c(8,8), #key="False",
#          Rowv=T, Colv=F, 
#          colsep = 1, 
#          offsetCol = 0, offsetRow = 0, #lhei=c(1,5), lwid = c(1,6),
#          linecol =  "black",
#          density.info = "none", trace = "none", dendrogram = "none",
#          cexRow = 0.75, cexCol = 0.75)
#dev.off()

```

* Note: Not depicted, is an analyses of phenotypic correlations of diet-confounder associations: We saw that adjusting for dPCs attenuates the associations modestly, although not completely, suggesting there are unmeasured confounders or mediators.


### Examine the effect of diet PC adjustment on causal effect estimates for variant-Food intake associations 
#### Examine the magnitude (%) change in Beta coefficients with dietPC adjustment

```{r}

# ==========================================================================
## Dot plot BETA (% change) when adjusting for EACH covariate
### Confounders and dietPCs
# ==========================================================================

plot_Bchange.fun <- function(Bchange, food, covars) {
  
  tab <- Bchange %>%
    pivot_Bdat_to_long() %>%
    mutate(Covar = factor(Covar, levels=c("Unadjusted", "DietPCs", 
                                          covarSets$confounders$Names, "All non-DietPC \nConfounders", 
                                          covarSets$dietpctop23$Names)))
  
  yscale <- ceiling(max(abs(tab$B_pctChange))*1.05)
  
  if(food == "raw_veg") { 
    food_colors = rev(palettes$classicgreen)
    } else { food_colors = rev(palettes$classicorange) }
  
  if(covars == "all") {
    plot_colors = c(palettes$Conf[1:7], food_colors)
  } else if(covars == "dPCs") {
    plot_colors = food_colors
  } else if (covars =="conf") {
    plot_colors = c(food_colors[1], palettes$Conf[1:7], "grey38")
  }

  tab %>%
    filter(Covar != "Unadjusted") %>%
    ggplot(aes(x=Covar, y = B_pctChange, color = Covar, group=snp)) + 
    theme_bw() + ggtheme2 +
    geom_hline(yintercept = 0, color = "black") + 
    geom_point(size=1.35, position=position_jitter(0.15), alpha=0.95) + 
    scale_y_continuous(limits=c(-yscale, yscale), breaks=c(-60, -40, -20, -10, 10, 20, 40, 60)) +
    geom_hline(yintercept = c(-10, 10), color = "black", linetype = "dashed") + 
    scale_color_manual(values =  plot_colors, name = "Confounder", guide=F) +
    #ylab("% change in SNP-diet estimate \n with confounder adjustment") + xlab(" ") + 
    ylab("% change in SNP-diet estimate") + xlab(" ") + 
    theme(axis.text.x = element_text(angle=25, hjust=1, size=8),
          axis.title = element_text(size=8), axis.text.y=element_text(size=8))
}


# ======================
## Make plots of: Effect of covariate adjustment on variant-diet estimates \n
# ======================

veg_Bchange.df <- rbind.data.frame(
  veg_Bdat.l$each_conf, 
  veg_Bdat.l$all_conf %>% mutate(Covar=gsub("Standard", "All non-DietPC \nConfounders", Covar)),
  veg_Bdat.l$all_dPCs %>% mutate(Covar=gsub("Top23DietPCs", "DietPCs", Covar)))

plot_Bchange_conf <- plot_Bchange.fun(veg_Bchange.df, "raw_veg", covars = "conf") 

#plot_Bchange_conf %>%  ggsave(filename = "../output/plot_raw_veg_v2_Bchange_conf.pdf", height = 4, width=5)
plot_Bchange_conf

```

In the plots of magnitude (%) Beta change after adjusting for hypothesized confounders/mediators, we see most substantial impact on Betas when adjusting for INCOME relative to other behavioral/lifestyle traits.


##### Examine the adjusted Beta coefficients after dietPC adjustment 
Diet PCs were progressively introduced into the model and the resulting beta (5% CI) were plotted

```{r fn-rediMR-adjB, fig.asp=0.55}
# ==========================================================================
## Function to plot BETA (adjusted) when ADDING adjustment for all dietPCs
# ==========================================================================

# Dot plot of pctBchange when adjusting for ALL covariates
plot_adjB_dPCall.fun<-function(food, Bchange, PCs_to_plot="all") {
  tab_all <- veg_Bchange.l$All_combined %>%
    pivot_longer(c(ends_with("base"), ends_with("adj")), names_to=c("stat", "mod"), names_sep="_") %>%
    pivot_wider(names_from="stat", values_from = value) %>%
    mutate(nPCs = ifelse(mod == "base", 0, nPCs)) %>%
    filter((Covar=="Top1DietPCs" & mod == "base") | mod == "adj") %>%
    mutate(shape=ifelse(nPCs==0, "square", "dot"))
  
  yscale<-max(abs(tab_all$lowCI*1.0), abs(tab_all$upCI*1.0))
  
  # If plotPCs != all, which PCs should be plotted?
  if(PCs_to_plot[1] == "all") {
    plotPCs <- 0:23} else{
      plotPCs <- c(0, PCs_to_plot)
    }
  
  # Set color palette
  if(food=="raw_veg") {
    plot_colors<-rev(palettes$classicgreen)[plotPCs]} else{
      plot_colors<-rev(palettes$classicorange)[plotPCs]
    } 
  
  tab_all %>% 
    filter(nPCs %in% plotPCs) %>%
    ggplot(aes(x=as.numeric(nPCs), y = B, group=snp, color = as.factor(nPCs), shape=shape)) + 
    theme_bw() + ggtheme2 + 
    geom_hline(yintercept = 0, color = "black") + 
    geom_point(size=1.55, alpha=0.95) +
    geom_line(linewidth=0.15) +
    scale_color_manual(values=c("#00000065", plot_colors)) +
    ylim(-yscale, yscale) +
    scale_x_continuous(breaks=plotPCs) +
    scale_shape_manual(values=c("square"=15, "dot"=19)) +
    #ylab("SNP-diet estimate (95% CI) \n with DietPC adjustment") + 
    ylab("SNP-diet estimates (95% CI)") + 
   # xlab("n dietPCs adjusted") +
    xlab(NULL) +
    theme(legend.position = "none",
          axis.text = element_text(size=8), axis.title = element_text(size=8))
}

# ======================
## Print & save plots 
# ======================

plot_adjB_dPC <- plot_adjB_dPCall.fun("raw_veg", veg_Bchange.l$All_combined) 

#plot_adjB_dPC %>% ggsave(filename="../output/raw_veg_v2_plot_adjB_dPCall.pdf", height = 4, width =6)
plot_adjB_dPC
```

#### Focus on adjusted Beta coefficients for each confounder, in referent models either without (Base GWAS covariates) or with (DietPCs) prior adjustment for diet PCs

```{r}

# ====================================
# Plot Absolute changes in Beta
# ====================================

tab_Bchange_forPlot <- rbind.data.frame(
  veg_Bdat.l$each_conf, veg_Bdat.l$each_dPCplusConf, 
  veg_Bdat.l$all_dPCs, veg_Bdat.l$all_conf,
  veg_Bdat.l$all_dPCplusConf %>% filter(Covar =="DietPCs_Add7cWaist2Hip")) %>% 
  left_join(veg_Bdat.l$all_dPCs %>% select(c("snp", B_dPC="B_adj")), by = "snp") %>%
  mutate(B_diff=B_adj-B_base) %>%
  pivot_Bdat_to_long() %>%
  mutate(B_diff = ifelse(Covar=="Unadjusted", 0, B_diff),
         Reference_Model = ifelse(startsWith(Covar, "DietPCs_"), "DietPC", "Base")) %>%
  mutate(B_diff=ifelse(startsWith(Covar, "DietPCs_")==T, B-B_dPC, B_diff)) %>%
  mutate(Covar = factor(Covar, levels=c("Unadjusted", "DietPCs", "Smoking", "DietPCs_Smoke", "Alcohol", "DietPCs_Alch", "Physical Activity", "DietPCs_PhysAct",
                                        "Income", "DietPCs_Income", "Education", "DietPCs_Educ", "BMI", "DietPCs_BMI", "Waist-to-hip", "DietPCs_Waist2Hip",
                                         "Standard", "DietPCs_Add7cWaist2Hip"),
                        labels=c("Unadjusted", "DietPCs", "Smoking", "DietPCs+Smoking", "Alcohol", "DietPCs+Alch", "Phys Act", "DietPCs+Phys Act",
                                 "Income", "DietPCs+Income", "Education", "DietPCs+Education", "BMI", "DietPCs+BMI", "Waist2Hip", "DietPCs+Waist2Hip",
                                "All non-DietPC\nConfounders", "DietPCs+All non-DietPC\nConfounders"))) %>%
  mutate(Covar_cat = gsub("DietPCs[+]", "", Covar)) %>%
  mutate(Covar_cat = ifelse(Covar_cat == "Alch", "Alcohol",ifelse(Covar_cat == "Add7cWaist2Hip", "All non-DietPC\nConfounders", Covar_cat))) %>%
  mutate(Covar_cat = factor(Covar_cat, levels=c("Unadjusted", "DietPCs", "Smoking", "Alcohol", "Phys Act", "Income", "Education", "BMI", "Waist2Hip", "All non-DietPC\nConfounders"))) %>%
  filter(complete.cases(Covar_cat))

plot_Bchange_abs_compare <- tab_Bchange_forPlot %>%
  filter(Covar_cat != "Unadjusted" & Covar != "DietPCs") %>%
  ggplot(aes(x=Reference_Model, y = B_diff, color = Covar, group=snp, shape=Reference_Model)) + 
  theme_bw() + ggtheme2 +
   facet_grid(~Covar_cat, scales = "free_x") +
  geom_hline(yintercept = 0, color = "black") + 
  geom_point(size=1.55, alpha=0.95) + 
  geom_line(linewidth=0.15) +
  scale_y_continuous(limits = c(-0.011, 0.01)) +
  scale_color_manual(values =  c(rep(palettes$Conf[1:7], each=2), "grey38", "grey38"), guide=F) +
  #ylab("Change in SNP-diet estimate from \nBase or DietPC referent model") + 
  ylab("Change in SNP-diet estimate from referent model") + 
  xlab("Referent Model") + #Effect of covariate adjustment
  theme(legend.position = "none", strip.text.x = element_text(size=8),
        axis.text = element_text(size=8),
        axis.title.y = element_text(size=8)) 

# Save
plot_Bchange_abs_compare #%>% ggsave(filename="../output/raw_veg_v2_plot_panel_Fig2_Bchange_abs_compare.pdf", height = 3, width =8)


```

### Pull out individual SNP examples where hypothesis is supported and not

First, focus on Income associations given the apparent heterogeneity in variant-diet associations

```{r}
# Focusing on Income and PA, which had most evidence of confounding based on change-in-estimate analysis;
# Visualizing the SNPs for which adjusting for Covariates ON TOP of dietPCs had the LEAST impact on effect estimates
# and the MOST impact on effect estimates

tab_Bchange_forPlot %>% filter(Reference_Model == "DietPC") %>% 
  mutate(B_pctChange_fromDiet = ((B-B_dPC)/B_dPC)*100, .before="B_pctChange") %>%
  filter(Covar_cat == "Phys Act") %>% arrange(B_pctChange_fromDiet)

# Zoom in on individual SNPs
plot_snp_adjB.fun <- function(variant, covar, with_labels=F) { 
  
  tab_plot <- tab_Bchange_forPlot %>% filter(snp == variant) %>%
    filter(Covar %in% c("Unadjusted", covar, "DietPCs", paste0("DietPCs+", covar))) %>%
    mutate(order=factor(Covar, levels=c("Unadjusted", covar, "DietPCs", paste0("DietPCs+", covar)))) %>%
    mutate(shape=ifelse(Covar == "Unadjusted" | Covar == covar, "refBase", "refDiet")) %>%
    mutate(labs=ifelse(Covar_cat==covar, round(B_diff, 3), NA))
  
  labs.y <-c(mean(tab_plot$B[tab_plot$shape == "refBase"]), mean(tab_plot$B[tab_plot$shape == "refDiet"]))
    
  ymax<-max(c(abs(tab_plot$upCI), abs(tab_plot$lowCI)))*1.05
  if(tab_plot$B[1]>0) {yscale <- c(0, ymax)} else{yscale <- c(-ymax, 0)}
  
  plot_colors <- c(palettes$ConfNames, "DietPCs"=rev(palettes$classicgreen)[1], palettes$ConfNames[covar])
  names(plot_colors)[10]=paste0("DietPCs+", covar)
  
  tab_plot <- tab_plot %>% mutate(labs.y = rep(labs.y, each=2))
  
  p<-tab_plot %>%
    ggplot(aes(x=order, y=B, ymin=lowCI, ymax=upCI, fill=order, color=order, shape=shape, group=shape)) + 
    #geom_bar(stat="identity", width=0.55) + 
    geom_point(size=3.5) +
    geom_line(linetype="dashed", color="black", linewidth=0.35) +
    geom_errorbar(width=0.25, linewidth=0.75) + 
    #geom_hline(yintercept = 0) + 
    ylab("") + xlab("") + 
    #scale_y_continuous(limits=yscale) +
    scale_fill_manual(values=plot_colors, guide=F) +
    scale_color_manual(values=plot_colors, guide=F) +
    ggtheme2 + 
    theme(legend.position = "none", title = element_text(face="bold"),
          axis.text.x = element_text(angle=25, hjust=1, size=8),
          axis.text.y = element_text(size=8)) +
    ggtitle(paste0(variant))
  
  if(with_labels == T) {
    p <- p +
      ylab("SNP-diet estimates (95% CI)") + xlab("") #+ 
      #geom_text(aes(x=order, y=labs.y+abs(labs.y)*.05, label=labs), color="black") + 
     # theme(text = element_text(face="bold", size=8))
  }
  return(p)
}

```

Then, expand to looking across all confounders

```{r}
# Gather SNP extremes: 
tab_Bchange_forSingleSNPs <- tab_Bchange_forPlot %>% 
  select(snp, Covar_cat, B_diff, B, Reference_Model, B_pctChange) %>%
  pivot_wider(values_from=c("B_diff"), names_from = "Reference_Model", id_cols = c(snp, Covar_cat)) %>%
  mutate(diff_of_diff = DietPC-Base, dietpc_to_base = DietPC/Base) %>% arrange(abs(dietpc_to_base))

#tab_Bchange_forSingleSNPs %>% filter(Covar_cat %in% c("Income"))

## Comparing (RELATIVE) changes in DietPC model to changes in Base model:
# SMALLEST change in DietPC relative to Base model: rs7834671_G (Income) ; rs67497633_G (PA)
p_snpxinc_supp_rs783 <- plot_snp_adjB.fun("rs7834671_G", "Income", with_labels = T) #; plot_snp_adjB.fun("rs7619139_T", "Income")
p_snpxinc_supp_rs761 <- plot_snp_adjB.fun("rs7619139_T", "Income") #; plot_snp_adjB.fun("rs7619139_T", "Income")
#p_snpxpa_supp <- plot_snp_adjB.fun("rs67497633_G", "Phys Act") #; plot_snp_adjB.fun("rs3828120_T", "Phys Act")

## Comparing (ABSOLUTE) change in dietPC model to change in Base model:
#tab_Bchange_forSingleSNPs %>% filter(Covar_cat %in% c("Income", "Phys Act")) %>% arrange(DietPC)

# LARGEST residual change in dietPC model
p_snpxinc_not_rs114 <- plot_snp_adjB.fun("rs114595701_C", "Income", with_labels = T)
p_snpxinc_not_rs775 <- plot_snp_adjB.fun("rs7753228_C", "Income")

plot_Bchange_snps <- ggarrange(
  ggarrange(p_snpxinc_supp_rs783, p_snpxinc_supp_rs761, ncol=2), "",
  ggarrange(p_snpxinc_not_rs114, p_snpxinc_not_rs775, ncol=2),
  ncol=3, widths = c(1, 0.1, 1))

plot_Bchange_snps
```

```{r, include=F}
# ====================================
# Combined panel plot
# ====================================
ggarrange(
  ggarrange(plot_adjB_dPC, "", plot_Bchange_conf, ncol=3, nrow=1, align="hv", widths=c(1.5, 0.05, 1)),
  plot_Bchange_abs_compare, plot_Bchange_snps,
  ncol=1, nrow=3) #%>% ggsave(filename="../output/raw_veg_v2_plot_panel_Fig2_v3.pdf", height=8 , width=10)

# ===================================================
## SNPs with the most and least DIETPC confounding
# ===================================================

veg_Bdat.l$all_dPCs %>%
  rowwise() %>% mutate(B_diff = B_adj - B_base) %>% rowwise() %>%
  arrange(abs(B_diff))

snps_most_dPC_conf <- c("rs62106252_G", "rs1269454_G")
snps_least_dPC_conf <- c("rs11209780_G", "rs67497633_G")


tab_Bchange_forPlot %>% filter(snp %in% c(snps_most_dPC_conf, snps_least_dPC_conf)) %>%
  filter(Reference_Model == "DietPC" | Covar == "Unadjusted" | Covar == "DietPCs") %>%
  ggplot(aes(y=B, ymin=lowCI, ymax=upCI, x=snp, group=Covar, color=Covar_cat)) + 
  geom_point(size=2.55, position=position_dodge(0.65)) + #geom_line() + 
  geom_errorbar(width=0.55, position=position_dodge(0.65)) + #geom_line() + 
  scale_color_manual(values=c(palettes$ConfNames, "DietPCs"=palettes$classicgreen[15],
                              "All non-DietPC\nConfounders"="brown")) + 
  geom_hline(yintercept = 0) + 
  ggtheme2 + 
  ylab("Beta (95% CI)") +
  theme(axis.text.x=element_text(angle=15, hjust=0.5, vjust=0.5))
  

```



```{r, include=F}

veg_Bdat.l$each_allvars %>% 
  filter(Covar %in% c("Diet Pattern PC23", covarSets$confounders$Names)) %>%
  mutate(snp=gsub("snp","",gsub("[.]",":",snp))) %>%
  mutate(across(is.numeric, ~round(., 4))) %>%
  mutate(Raw_B.SE = sprintf("%s (%s)", B_base, SE_base),
         Conf_B.SE = sprintf("%s (%s)", B_adj, SE_adj)) %>%
  select(snp, Covar, Raw_B.SE, Conf_B.SE) %>%
  pivot_wider(values_from = c("Raw_B.SE", "Conf_B.SE"), names_from="Covar") %>%
  rename("Unadjusted_B.SE" = "Raw_B.SE_Smoking") %>%
  select(-starts_with("Raw")) #%>%
  #write.csv("../output/raw_veg_v2_tab_BadjbyConf.csv")


## Take variants that change at the extremes
veg_Bchange.l$All$dietpctop23[[1]] %>% arrange(B_pctChange)
veg_snps_extreme <- c("rs1269454_G", "rs2941732_A", "snp2.104130139_AT_A_AT", "rs3096690_G", "rs11209780_G", "rs67497633_G")

fruit_Bchange.l$All$dietpctop23[[1]] %>% arrange(B_pctChange)
fruit_snps_extreme <- c("rs429358_T", "rs62106258_T", "rs35628463_G", "rs2081194_G", "rs1052486_A", "rs144893360_T")

```


```{r, include=F}
# ========================================================================
## ZOOMED IN plot of BETA (adjusted) when adjusting for each CONFOUNDER
# ========================================================================

plot_adjB_confZoom.fun <- function(Bdat.l, snp_extremes) {
  
  tab <- rbind.data.frame(veg_Bdat.l$each_conf, veg_Bdat.l$all_dPCs, veg_Bdat.l$all_conf, 
                          veg_BchangeAddConf[[7]]) %>%
    filter(snp %in% veg_snps_extreme) %>%
    pivot_longer(c(ends_with("base"), ends_with("adj")), names_to=c("stat", "mod"), names_sep="_") %>%
    pivot_wider(names_from="stat", values_from = value) %>%
    mutate(Covar = ifelse(Covar == "Smoking" & mod == "base", "Unadjusted", 
                          ifelse(Covar == "Standard", "All_Confounders", 
                                 ifelse(Covar == "DietPCs_Add7cWaist2Hip", "DietPCs + All Confounders", Covar))),
           snp=gsub("snp","",gsub("[.]",":", snp))) %>%
    filter(Covar == "Unadjusted" | mod == "adj") %>% 
    mutate(Covar = factor(Covar, levels=c("Unadjusted", covarSets$confounders$Names, 
                                          "All_Confounders", "DietPCs", "DietPCs + All Confounders"))) %>%
    mutate(B_pctChange = ifelse(Covar == "Unadjusted", 0, B_pctChange)) 
  
  yscale<-c(min(tab$lowCI)*1.05, max(tab$upCI)*1.05)
  
  tab %>%
    ggplot(aes(x=snp, y = B, ymin=lowCI, ymax=upCI, color = Covar)) + 
    theme_bw() + ggtheme2 +
    theme(axis.text.x = element_text(angle=35, hjust=1, size=8),
          panel.grid.major.x = element_line(),
          legend.position = "bottom") +
    geom_point(size=1.15, position=position_dodge(0.65), alpha=0.95) + 
    geom_errorbar(width=0.55, position=position_dodge(0.65), alpha=0.95, linewidth=0.35) + 
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") + 
    scale_y_continuous(limits=c(yscale[1], yscale[2])) +
    scale_color_manual(values =  c("black", palettes$ConfNames[1:7], "brown", 
                                   rev(palettes$classicgreen)[1], "darkgrey"), name = "Adjustments") +
    ylab("Beta (95% CI)") + xlab(" ") 
}

plot_adjB_confZoom.fun(veg_Bchange_conf.l, veg_snps_extreme)

```


```{r, width=12}

plot_adjB_confZoom2.fun <- function(Bdat.l, BchangeEachConf, snp_extremes) {
  
  tab <- rbind.data.frame(Bdat.l$each_conf, Bdat.l$all_dPCs, Bdat.l$all_conf, 
                          do.call(rbind.data.frame, BchangeEachConf) %>%
                            mutate(Covar=rep(paste0("DietPCs+", covarSets$confounders$Names), each=length(unique(snp)) )) 
                          ) %>%
    filter(snp %in% veg_snps_extreme) %>%
    pivot_longer(c(ends_with("base"), ends_with("adj")), names_to=c("stat", "mod"), names_sep="_") %>%
    pivot_wider(names_from="stat", values_from = value) %>%
    mutate(Covar = ifelse(Covar == "Smoking" & mod == "base", "Unadjusted", 
                          ifelse(Covar == "Standard", "Confounders", Covar)),
           snp=gsub("snp","",gsub("[.]",":", snp))) %>%
    mutate(ifelse(is.na(Covar)==T, "Waist2Hip", Covar)) %>%
    filter(Covar == "Unadjusted" | mod == "adj") %>% 
    mutate(B_pctChange = ifelse(Covar == "Unadjusted", 0, B_pctChange)) %>%
    mutate(Covar2=factor(ifelse(Covar == "Top23DietPCs", "DietPCs", Covar),
                         levels=c("Unadjusted", covarSets$confounders$Names, "Confounders", "DietPCs", paste0("DietPCs+", covarSets$confounders$Names)))
           )

  yscale<-c(min(tab$lowCI)*1.05, max(tab$upCI)*1.05)
  
  tab %>%
    ggplot(aes(x=snp, y = B, ymin=lowCI, ymax=upCI, color = Covar2)) + 
    theme_bw() + ggtheme2 +
    theme(axis.text.x = element_text(angle=35, hjust=1, size=8),
          panel.grid.major.x = element_line(),
          legend.position = "bottom") +
    geom_point(size=1.15, position=position_dodge(0.65), alpha=0.95) + 
    geom_errorbar(width=0.55, position=position_dodge(0.65), alpha=0.95, linewidth=0.35) + 
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") + 
    scale_y_continuous(limits=c(yscale[1], yscale[2])) +
    scale_color_manual(values =  c("black", palettes$Conf[1:7], "brown", 
                                   rev(palettes$classicgreen)[1], palettes$Conf[1:7]), name = "Adjustments") +
    ylab("Beta (95% CI)") + xlab(" ") 
}

plot_adjB_confZoom2.fun(veg_Bdat.l, veg_BchangeEachConf, veg_snps_extreme)

```


##### Summary:
- Note: adjusting for diet PCs attenuated all SNP-diet associations, but not to 0, indicating (1) remaining/unmeasured confounders/mediators explaining SNP-diet relations, or (2) possible direct effects of the SNP on diet, not explained by behavioral/lifestyle confounding or mediation.

- We then compared the effects of dPC adjustment to confounder adjusment. We hypothesized that adjusting for each confounder will attenuate the associations to a modest degree; and that adjusting for an appropriate amount of dPCs will attenuate the associations to a greater degree, representing correction for behaviora/lifestyle confounders. 



